---
title: "Video Classification"
author: "Gustav Idun Sloth"
date: "2023-11-18"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
```

```{r loading data, warning = FALSE}

setwd(".")
df <- read_csv("data/video_dataset/dataset_10k_lang_5.0.csv")

```

## Data Preperation

We remove all non-English entries as well as unnecessary columns and normalize row naming conventions.

```{r data prep}

df <- df %>%
  select(title:language) %>%
  filter(language == "en") %>%
  rename(view_count = viewCount) %>%
  drop_na(view_count)

nrow(df)
  
```

## Plotting Distributions

We see that the view_count distribution is heavily imbalanced. 

```{r distribution plotting}

plot_dist <- function(data){
  quantiles <- quantile(data$view_count, probs = c(0,.10, .20, .30, .40, .50, .60, .70, .80, .90, 1))
  print(quantiles %>% as.data.frame())
  
  ggplot(data, aes(x=view_count)) +
    geom_histogram(binwidth = 0.1) +
    scale_x_log10(labels = label_comma()) +
    scale_y_continuous(labels = label_comma()) +
    labs(title = "View Count Distribution", x = "View Count (log10 scaled)", y = "Count") +
    geom_vline(xintercept = quantiles) +
    theme_minimal()
}

plot_dist(df)

```

## Defining Classes and Resampling data

First we define three classification classes that our model will use. 

Then we create both an unbalanced and a balanced dataset using undersampling for the majority class, "medium".

```{r imbalanced classification}

df_classes <- list(
  "low" = df %>% filter(view_count <= 1000) %>% mutate(classification = 0),
  "medium" = df %>% filter(view_count > 1000 & view_count < 10000) %>% mutate(classification = 1),
  "high" = df %>% filter(view_count >= 10000) %>% mutate(classification = 2)
)

df_unbalanced <- bind_rows(df_classes)

display_class_nrow <- function(){
  
  nrows <- list(
    "low" = nrow(df_classes$low),
    "medium" = nrow(df_classes$medium),
    "high" = nrow(df_classes$high)
  )
  
  total_rows = sum(unlist(nrows))
  
  tibble("Low (<1.000)" = c(nrows$low, round(nrows$low/total_rows*100)), "Medium (1.000-10.000)" = c(nrows$medium, round(nrows$medium/total_rows*100)), "High (>10.000)" = c(nrows$high, round(nrows$high/total_rows*100)))

}

percentile <- ecdf(df$view_count)
percentile(1000)
percentile(10000)


display_class_nrow()

df_classes$medium <- sample_frac(df_classes$medium, size = 0.54)

df_balanced <- bind_rows(df_classes)

display_class_nrow()

print(paste0("n rows after undersampling: ", nrow(df_balanced)))

plot_dist(df_balanced)

```

## Write Datasets to Files

```{r save data}

save_and_shuffle <- function(data, filename){
  data_shuffled <- data[sample(nrow(data)),]
  write_csv(data_shuffled, paste0(filename, ".csv"))
}

save_and_shuffle(df_unbalanced, "model_dataset_unbalanced")
save_and_shuffle(df_balanced, "model_dataset_balanced")

```